Supplementary material for\\ A Benchmark of Named Entity Recognition Approaches in Historical Documents\\ Application to 19$^{th}$ Century French Directories

### 2022-01-20 src-latex/parts/20-related-work.tex
Fixed quirks in section 5

Based on those works,

\subsection{== À réintégrer depuis ancienne section 4 ==}

A introduire en disant quo'n chercher à utiliser au plus des modèles simples à reprendre: déjà entraînés, dispos pour du français et adaptables
We select two deep-learning-based NER models available in packaged software libraries: SpaCy NLP pipelines and CamemBERT.

>>>>

Based on those works, we adopt two French deep-learning-based language models, available in packaged software libraries, already trained for the NER task and that we can adapt to the domain of historical directories: SpaCy NLP pipelines and CamemBERT.

### 2022-01-20 src-latex/parts/50-ner-xp1-training-set-size-and-pretraining.tex
Fixed quirks in section 5

Then, a stratified sampling based on the source directory of each entry is run on the remainder to create a training set (6373 entries, $\approx 73\%$ of the gold reference) and a development set (709 entries, $\approx 8\%$).
The development set is used to evaluate the model performance during the training phase.
This resampling procedure is a convenient way to shape both sets, so they reflect the diversity of directories within the ground truth.

@@

The three models are trained 5 times on each of the 8 training subsets using an early stopping criterion based on the number of training steps without improvement of the F1-score. 
This patience threshold is set at 1600 steps for SpaCY NER, and 3 evaluations (1 evaluation every 100 steps) for CmBERT and CmBERT+ptrn.
The metrics are measured for the 24 resulting NER models on the common test and averaged over the runs.

@@

\begin{table}[h!]
\centering
\caption{F1 score, precision and recall measured on the fine-tuned models CmBERT, CmBERT+ptrn and SpaCy NER in experiment 1 for 8 training sets of increasing sizes.}
\begin{tabular}{llrrrrrrrr}
       & Training examples &  49   &  99   &  199  &  398  &  796  &  1593 &  3186 &  6373 \\
       & \% & 0.8   & 1.6   & 3.1   & 6.2   & 12.5  & 25.0  & 50.0  & 100.0 \\
\midrule\bottomrule
\multirow{3}{*}{\rotatebox{90}{F1 score}} & CmBERT &  89.5 &  90.5 &  92.7 &  93.3 &  \textbf{94.1} &  \textbf{94.9} &  \textbf{94.6} &  \textbf{95.1} \\
       & CmBERT-ptrn &  \textbf{92.2} &  \textbf{92.9} &  \textbf{93.6} &  \textbf{93.8} &  93.8 &  94.1 &  \textbf{94.6} &  94.4 \\
       & SpaCy NER &  87.0 &  89.0 &  90.3 &  91.9 &  92.1 &  92.8 &  93.2 &  93.5 \\
\cline{1-10}
\multirow{3}{*}{\rotatebox{90}{Precision}} & CmBERT &  87.4 &  88.7 &  91.5 &  92.7 &  93.3 &  94.9 &  93.9 &  95.1 \\
       & CmBERT-ptrn &  90.8 &  91.8 &  92.9 &  93.0 &  93.0 &  93.4 &  94.1 &  93.9 \\
       & SpaCy NER &  85.6 &  87.7 &  90.0 &  92.0 &  92.4 &  92.8 &  93.1 &  93.7 \\
\cline{1-10}
\multirow{3}{*}{\rotatebox{90}{Recall}} & CmBERT &  91.6 &  92.5 &  93.9 &  93.9 &  94.9 &  94.9 &  95.4 &  95.1 \\
       & CmBERT-ptrn &  93.6 &  94.0 &  94.4 &  94.6 &  94.6 &  94.8 &  95.0 &  94.9 \\
       & SpaCy NER &  88.6 &  90.4 &  90.7 &  91.7 &  91.9 &  92.8 &  93.3 &  93.4 \\
\end{tabular}
\label{tab:experiment-1-models-performances}
\end{table}

\begin{figure}[h!]
	   \center{\includegraphics[width=\textwidth]
	       {images/experiment-1-models-performances.pdf}}
	  \caption{\label{fig:f1-vs-trainsize} Graph view of the metrics displayed in ~\ref{tab:experiment-1-models-performances}. As SpaCy NER is outperformed by BERT models from the start, CmBERT slightly surpasses the pre-trained version when the number of training samples grows.}.
\end{figure}

>>>>

Then, a stratified sampling based on the source directory of each entry is run on the remainder to create a training set (6373 entries, $\approx 73\%$ of the gold reference) and a development set (709 entries, $\approx 8\%$).
The development set is used to evaluate the model during the training phase.
This resampling procedure is a convenient way to shape both sets, so they reflect the diversity of directories within the ground truth.

@@

The three models are fine-tuned on the NER task 5 times using each of the 8 training subsets, with an early stopping criterion based on the number of training steps without improvement of the F1-score. 
This patience threshold is set to 1600 steps for SpaCY NER and 3 evaluations (1 evaluation every 100 steps) for CmBERT and CmBERT+ptrn.
The metrics are measured for the 24 resulting NER models on the common test set and averaged over the runs.

@@

\begin{table}[h!]
\centering
\caption{\label{tab:experiment-1-models-performances} F1 score, precision and recall measured on the fine-tuned models CmBERT, CmBERT+ptrn and SpaCy NER in experiment 1 for 8 training sets of increasing sizes.}
\begin{tabular}{llrrrrrrrr}
       & Training examples &  49   &  99   &  199  &  398  &  796  &  1593 &  3186 &  6373 \\
       & \% & 0.8   & 1.6   & 3.1   & 6.2   & 12.5  & 25.0  & 50.0  & 100.0 \\
\midrule\bottomrule
\multirow{3}{*}{\rotatebox{90}{F1 score}} & CmBERT &  89.5 &  90.5 &  92.7 &  93.3 &  \textbf{94.1} &  \textbf{94.9} &  \textbf{94.6} &  \textbf{95.1} \\
       & CmBERT-ptrn &  \textbf{92.2} &  \textbf{92.9} &  \textbf{93.6} &  \textbf{93.8} &  93.8 &  94.1 &  \textbf{94.6} &  94.4 \\
       & SpaCy NER &  87.0 &  89.0 &  90.3 &  91.9 &  92.1 &  92.8 &  93.2 &  93.5 \\
\cline{1-10}
\multirow{3}{*}{\rotatebox{90}{Precision}} & CmBERT &  87.4 &  88.7 &  91.5 &  92.7 &  93.3 &  94.9 &  93.9 &  95.1 \\
       & CmBERT-ptrn &  90.8 &  91.8 &  92.9 &  93.0 &  93.0 &  93.4 &  94.1 &  93.9 \\
       & SpaCy NER &  85.6 &  87.7 &  90.0 &  92.0 &  92.4 &  92.8 &  93.1 &  93.7 \\
\cline{1-10}
\multirow{3}{*}{\rotatebox{90}{Recall}} & CmBERT &  91.6 &  92.5 &  93.9 &  93.9 &  94.9 &  94.9 &  95.4 &  95.1 \\
       & CmBERT-ptrn &  93.6 &  94.0 &  94.4 &  94.6 &  94.6 &  94.8 &  95.0 &  94.9 \\
       & SpaCy NER &  88.6 &  90.4 &  90.7 &  91.7 &  91.9 &  92.8 &  93.3 &  93.4 \\
\end{tabular}
\end{table}

\begin{figure}[h!]
	   \center{\includegraphics[width=\textwidth]
	       {images/experiment-1-models-performances.pdf}}
	  \caption{\label{fig:f1-vs-trainsize} Graph view of the metrics displayed in~\cref{tab:experiment-1-models-performances}. As SpaCy NER is outperformed by both BERT models from the start, CmBERT slightly surpasses the pre-trained version when the number of training examples becomes large.}.
\end{figure}
